---
title: "R Notebook"
output:
  html_notebook: default
  html_document:
    toc: true
    toc_depth: 2
---

# Installation

* We use renv package for package version management and git for version control. Call renv::snapshot() to save state call renv::restore() to revert to the previous state.
* We install devtools, the last version of sparklyr and some data.

```{r}
install.packages(c("devtools", "tidyverse"))
devtools::install_github("rstudio/sparklyr")
install.packages(c("nycflights13", "Lahman"))
library(sparklyr)
library(tidyverse)
spark_available_versions()
spark_uninstall(version = '3.0.1', hadoop_version = '3.2')
spark_install(version = '3.0.1', hadoop_version = '3.2')
spark_installed_versions()
renv::snapshot()
```

We connect to spark and copy some data from R into the Spark:

```{r}

sc <- spark_connect(master = "local")
iris_tbl <- copy_to(sc, iris)
flights_tbl <- copy_to(sc, nycflights13::flights, "flights")
batting_tbl <- copy_to(sc, Lahman::Batting, "batting")
src_tbls(sc)

```

We can filter data in Spark:

```{r}
flights_tbl %>% filter(dep_delay == 2)
```

And also do some aggregation and plotting:

```{r}
delay <- flights_tbl %>%
  group_by(tailnum) %>%
  summarise(count = n(), dist = mean(distance), delay = mean(arr_delay, na.rm = TRUE)) %>%
  filter(count > 20, dist < 2000, !is.na(delay)) %>%
  collect
ggplot(delay, aes(dist, delay)) +
  geom_point(aes(size = count), alpha = 1/2) +
  geom_smooth() +
  scale_size_area(max_size = 2)
```

Where an aggregation function, like sum() and mean(), takes n inputs and return a single value, a window function returns n values. The output of a window function depends on all its input values, so window functions don’t include functions that work element-wise, like + or round(). Window functions include variations on aggregate functions, like cumsum() and cummean(), functions for ranking and ordering, like rank(), and functions for taking offsets, like lead() and lag().
 
```{r}
batting_tbl %>%
  select(playerID, yearID, teamID, G, AB:H) %>%
  group_by(playerID) %>%
  filter(min_rank(desc(H)) <= 2 & H > 0) %>%
  arrange(playerID, yearID, teamID) 

```

Using SQL to query Spark

```{r}
library(DBI)
iris_preview <- dbGetQuery(sc, "SELECT * FROM iris LIMIT 10")
iris_preview
```



